Rroject for the "Getting-and-Cleaning-Data" Course
===================================

*** 
#### **Assignment** 

* **Data: **

    + The data collected from the accelerometers from the Samsung Galaxy S smartphone. A full description is available at the site where the data was obtained: http://archive.ics.uci.edu/ml/datasets/Human+Activity+Recognition+Using+Smartphones
    + The data for the project: https://d396qusza40orc.cloudfront.net/getdata%2Fprojectfiles%2FUCI%20HAR%20Dataset.zip

***
* **Task:**

    The R script called run_analysis.R will do the following tasks:

    + Merges the training and the test sets to create one data set.
    + Extracts only the measurements on the mean and standard deviation for each measurement.
    + Uses descriptive activity names to name the activities in the data set
    + Appropriately labels the data set with descriptive variable names.
    + From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

*** 
#### **The run_analysis.R script**

The script has five sections each of which deals with one of the tasks in the assignment.  

1. Merges the training and the test sets to create one data set.

Read file into dataset and combine

```{r}

setwd("C./UCI_HAR_Dataset")  # set the working directory
list.files(path = ".")  # check the folders and files under the directory
list.files(path = "./test")
list.files(path = "./train")
x_test <- read.table("./test/X_test.txt")
x_train <- read.table("./train/X_train.txt")
x <- rbind(x_train,x_test)
y_test <- read.table("./test/Y_test.txt")
y_train <- read.table("./train/Y_train.txt")
y <- rbind(y_train,y_test)
s_test <- read.table("./test/subject_test.txt")
s_train <- read.table("./train/subject_train.txt")
s <- rbind(s_train,s_test)
alldata <- cbind(s,y,x)
``` 

2. Extracts only the measurements on the mean and standard deviation for each measurement.

```{r}
ftr <- read.table("./features.txt")
ftr_mean <- ftr[grep(".*mean.*", ftr$V2, ignore.case=TRUE),]
ftr_sdv <- ftr[grep(".*std.*", ftr$V2, ignore.case=TRUE),]
ftr_sub <- rbind(ftr_mean,ftr_sdv)
```

The following code seems to work as well

```{r}
ftr_sub2 <- ftr[grep("std|mean", ftr$V2, ignore.case=TRUE),]
```

3. Uses descriptive activity names to name the activities in the data set.

```{r}
colnames(alldata)[1:2] <- c("subject", "activity.code")
act <- read.table("./activity_labels.txt")
head(act)
colnames(act) <- c("activity.code","activity")
act[, 2] = tolower(gsub("_", " ", as.character(act[, 2])))
alldata_n <- merge(alldata, act, by = "activity.code")
alldata_n2 <- cbind(alldata_n[,2],alldata_n[,564],alldata_n[,3:563])
```

4. Appropriately labels the data set with descriptive variable names.

```{r}
ftr[,2] <- as.character(ftr[,2])
vars <- ftr[,2]
colnames(alldata_n2) <- c("subject","activity",vars)
write.table(alldata_n2, "cln_mrg_data.txt")
```

5. From the data set in step 4, creates a second, independent tidy data set with the average of each variable for each activity and each subject.

```{r}
install.packages("reshape2")
library(reshape2)
alldata_n2$activity <- as.factor(alldata_n2$activity)
alldata_n2$subject <- as.factor(alldata_n2$subject)
data_melted <- melt(alldata_n2, id = c("subject", "activity"))
data_avg <- dcast(data_melted, subject + activity ~ variable, mean)
data_avg <- data_avg[order(data_avg$subject,data_avg$activity)]

write.table(data_avg, "average_data.txt")
```


*** 
#### **The average_data.txt File**

This file contains the cleaned data after Step 5.